\textbf{\sffamily\Huge \ShortTitle}\\ {\textbf{\sffamily\Large \FullTitle}} \vspace{1cm}

### 2015-01-08 proposal/sections/costs.tex
Costs
cost updates

\section{Costs}
\label{sect:costs}
\subsection{Person}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
Person & Commitment & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Project Manager & 100\% & 54000 & • & • & • & • \\ 
\hline 
Quality Manager & 75\% & 27000 & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
• & • & • & • & • & • & • \\ 
\hline 
\end{tabular}

\subsection{Hardware}
\begin{tabular}{|c|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Commercial EEG Headsets & 3000 & • & • & • & • \\ 
\hline 
Medical EEG Device & 60000 & • & • & • & • \\ 
\hline 
Server & 4000 & • & • & • & • \\ 
\hline 
Workstations & Notebooks & 6400 & • & • & • & • \\ 
\hline 
\end{tabular}
\subsection{Material}
\begin{tabular}{|c|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Hardware for the Prototype & 2000 & • & • & • & 2000 \\ 
\hline 
EEG Sensors & 500 & • & • & • & 500 \\ 
\hline 
• & • & • & • & • & • \\ 
\hline 
\end{tabular} 
\subsection{Travel Costs}
\begin{tabular}{|c|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Conferences & 4000 & • & 4000 & • & 8000 \\ 
\hline 
EEG Sensors & 500 & • & • & • & 500 \\ 
\hline 
• & • & • & • & • & • \\ 
\hline 
\end{tabular} 
\subsection{Other Costs}
\begin{tabular}{|c|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Teambuilding, Snacks, Coffee, Meetings & 7000 & 7000 & 7000 & 3500 & 8000 \\
\hline 
\end{tabular}

>>>>

\section{Costs}
\label{sect:costs}
\subsection{Personnel}

\begin{tabular}{|l|c|c|c|c|c|c|}
\hline 
Person & \% & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Project Manager & 100\% & 54000,00 & 55080,00 & 56181,60 & 28652,62 & 193914,22 \\ 
\hline 
Quality Manager & 75\% & 27000,00 & 27540,00 & 28090,80 & 14326,21 & 96957,11 \\ 
\hline 
Neuroscience Researcher 1 & 100\% & 41810,00 & 42646,20 & - & - & 84856,20 \\ 
\hline 
Neuroscience Researcher 2 & 100\% & 41810,00 & 42646,20 & - & - & 84856,20 \\ 
\hline 
Computer Science R. 1 & 100\% & 41810,00 & 42646,20 & 43499,12 & 22184,55 & 150139,88 \\ 
\hline 
Computer Science R. 2 & 100\% & 41810,00 & 42646,20 & 43499,12 & 22184,55 & 150139,88 \\ 
\hline 
Hardware Engineer 1 & 100\% & 31160,00 & 31783,20 & 32418,86 & 16533,62 & 111895,67 \\ 
\hline 
Hardware Engineer 2 & 100\% & 31160,00 & 31783,20 & 32418,86 & 16533,62 & 111895,67 \\ 
\hline 
Software Engineer 1 & 100\% & 31160,00 & 31783,20 & - & - & 62943,20 \\ 
\hline 
Software Engineer 2 & 100\% & 31160,00 & 31783,20 & - & - & 62943,20 \\ 
\hline 
\textbf{SUM} & & 372880,00 & 380337,60 & 236108,38 & 120415,27 & 989325,98 \\ 
\hline 
\end{tabular}

\paragraph{Project Manager}
A project manager supervises the project from the beginning to the end. His role has scheduling and supportive character, works close to everyone an thus has a good grasp of all parts of the project. Despite having a good overview, he is not deeply involved in any of the other employers work and not accumulating too specific knowledge. This way the risks involving a switch of the project managers position should theoretically be less of an impact then changing an expert.

\paragraph{Quality Manager}
The quality manager has a supportive work. Inspecting, improving and creating new standards during the whole project lifecycle helps raising work and product quality nd avoiding deadline misses. As the team is not very big, we think of a 75\% position %< sotz ausformulieren

\paragraph{Neuroscience Researcher}
The neuroscience researcher has a highly specialized research and develop task at the beginning of the project. To avoid risks and speed up the initial research two neuroscientists get hired.

\paragraph{Computer Science Researcher}
The computer science researchers care about algorithms and machine learning. They will be occupied  during the whole project. Two researcher should avoid risks of loosing knowledge if one leaves the project and can speed up the research especially within the first ans second milestone. Their field of work overlaps with all other project members.

\paragraph{Hardware Engineer}
The hardware engineer starts with analyzing existing devices and sensors. Working together with neuroscientists they create the Wavy prototype and are therefore part of the project during the whole 42 month. Two engineers provide a minimal backup.

\paragraph{Software Engineer}
The software engineers start implementing the knowledge gathered by neuroscience and computer science researcher starting with the second milestone. As for most positions two members seems fit to provide backup.

\subsection{Equipment}
\begin{tabular}{|l|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Commercial EEG Headsets & 3000,00 & - & - & - & 3000,00 \\ 
\hline 
Medical EEG Device & 60000,00 & - & - & - & 60000,00 \\ 
\hline 
Server & 4000 & - & - & - & 4000,00 \\ 
\hline 
Workstations/Notebooks & 6400 & - & - & - & 6400,00 \\ 
\hline 
\textbf{SUM} & 77000,00 &  &  &  & 77000,00 \\ 
\hline 
\end{tabular}

The equipment, which is needed for the WaveMeInProject will be purchased in the 1st year, since the equipment is required already in the reaserch part of the project and includes following hardware parts:
hardware Commercial EEG Headsets with costs of overall 3000\euro{} - the price per item is 499\euro{} and the project requires 6 pieces - https://emotiv.com/store/epoc-detail/

\subsection{Material}
\begin{tabular}{|l|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Hardware for the Prototype & 2000 & - & - & - & 2000 \\ 
\hline 
EEG Sensors & 500 & - & - & - & 500 \\ 
\hline 
\textbf{SUM} & 2500 &  &  &  & 2500 \\ 
\hline 
\end{tabular}

\subsection{Travel Costs}
\begin{tabular}{|l|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Conferences & 4000 & - & 4000 & - & 8000 \\ 
\hline 
\textbf{SUM} & 4000 &  & 4000 &  & 8000 \\ 
\hline 
\end{tabular}

The travel costs are estimated to be 8000\euro{}, whereof for the 1st and for the 3rd year 4000\euro{} are provided for advanced training courses.

\subsection{Other Costs}
\begin{tabular}{|l|c|c|c|c|c|}
\hline 
Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
\hline 
Teambuilding, Snacks, Coffee, Meetings & 7000 & 7000 & 7000 & 3500 & 24500 \\
\hline 
\end{tabular}

The other costs for teambuilding, snacks, coffee or meetings are estimated to be in sum 24500\euro{}: 7000\euro{} in the 1st year, 7000\euro{} in the 2nd year, 7000\euro{} in the 3rd year and 3500\euro{} in the 4th year.
